---
title: "Introduction: Time-series"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Time-series
A time-series is a chronological sequence of observations on a particular variable. For example the data below is a time-series which gives the quarterly total value of time deposits held by the Baarth National Bank during 1977 and 1978.

```{r echo = FALSE, results = 'asis'}
library('knitr')
library('kableExtra')
deposits <- data.frame('Quarter' = c('1977Q1','1977Q2','1977Q3','1977Q4','1978Q1','1978Q2','1978Q3','1978Q4'), 'Value' = c(35.3, 37.6, 38.1, 39.5, 37.9, 39.9, 40.1, 41.2))

knitr::kable(deposits, caption = 'Time Series Data: Quarterly Values of Time-Deposits') %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
# knitr::kable(deposits, caption = 'Time Series Data: Quarterly Values of Time-Deposits') 
```

Time-series data are often examined in hopes of discovering a historical pattern that can be exploited in the preparation of a forecast. In order to identify this pattern, it is often convenient to think of a time-series as consisting of several components:

* Trend
* Cycle
* Seasonal variations
* Irregular fluctuations

