---
title: "Introduction: Time-series"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Time-series
A time-series is a chronological sequence of observations on a particular variable. For example the data below is a time-series which gives the quarterly total value of time deposits held by the Baarth National Bank during 1977 and 1978.

```{r echo = FALSE, results = 'asis'}
library('knitr')
library('kableExtra')
deposits <- data.frame('Quarter' = c('1977Q1','1977Q2','1977Q3','1977Q4','1978Q1','1978Q2','1978Q3','1978Q4'), 'Value' = c(35.3, 37.6, 38.1, 39.5, 37.9, 39.9, 40.1, 41.2))

knitr::kable(deposits, caption = 'Time Series Data: Quarterly Values of Time-Deposits') %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
# knitr::kable(deposits, caption = 'Time Series Data: Quarterly Values of Time-Deposits') 
```

Time-series data are often examined in hopes of discovering a historical pattern that can be exploited in the preparation of a forecast. In order to identify this pattern, it is often convenient to think of a time-series as consisting of several components:

* Trend
* Cycle
* Seasonal variations
* Irregular fluctuations

*Trend* refers to the upward or downward movement that characterises a time-series over a period of time. Thus, trend reflects the long-run growth or decline in the time-series.  
Trend movements can represent a variety of factors, such as technological changes, changes in consumer tastes, increase in per capita income, changes in demographic, etc.  
*Cycle* refers to recurring up and down movements around trend levels. These fluctuations can have a duration of anywhere from 2 to 10 years or even longer measured from 'peak-to-peak' or trough-to-trough. One of the common cyclical fluctuations found in time-series data is the "business cycle", characterised by periods of prosperity and recession. There is no single explanation for cyclical fluctuations and they vary greatly in both length and magnitude.  
*Seasonal variations* are periodic patterns in a time-series that complete themselves within the period of a calendar year and are then repeated on a yearly basis. Seasonal variations are usually caused by factors such as weather and customs. Think, for example, the sales of on-line shops that is normally increasing during holiday season or think the temperature which is usually high during summer months and drops during winter.  
*Irregular fluctuations* are erratic movements in a time-series that follow no recognisable or regular pattern. Such movements represent what is "left over" in a time-series after trend, cycle and seasonal variations have been accounted for. Many irregular fluctuations in time-series are caused by "unusual" events that cannot be forecasted, such as extreme weather events, wars, strikes, etc. Irregular fluctuations can be also caused by errors on the part of the time-series analyst when, for example, entering data.  
Components discussed above do not always occur alone; they can occur in any combination or can occur all together. Based on the aforementioned there is no single best forecasting technique that fits all data. For example, a forecasting technique that is appropriate for a time-series characterised by trend alone may not be appropriate in forecasting time-series characterised by a combination of trend and seasonal variations.  
The graphs below represent 