---
title: "Time-series: A primer"
author: "themellion"
date: "8 December 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a very basic introduction into applied time-series using R, based on material coming from "Forecasting & time-series: An applied approach" by Bowerman & O'Connell. Topics that will be covered among others are: simple linear regression, exponential smoothing, ARIMA models, etc, while using R. 

## Time-series
A time-series is a chronological sequence of observations on a particular variable. For example the data below is a time-series which gives the quarterly total value of time deposits held by the Baarth National Bank during 1977 and 1978.

```{r echo = FALSE, results = 'asis}
deposits <- data.frame('Quarter' = c('1977Q1','1977Q2','1977Q3','1977Q4','1978Q1','1978Q2','1978Q3','1978Q4'), 'Value' = c(35.3, 37.6, 38.1, 39.5, 37.9, 39.9, 40.1, 41.2))

kable(deposits, caption = 'Time Series Data: Quarterly Values of Time-Deposits') 
```
